---
title: "Portfolio 2"
author: "Cassandra Richardson"
date: "4/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(tidyverse)
```

# **Cassandra Richardson's Portfolio Piece #2 - Data Analysis Report on Chicken Weight and Diet**



# Step 1: Finding a dataset.

### R has several built-in datasets that we can use to practice R functions with. To view these datasets, simply use the function below: 

```{r finding dataset}
data()
```

### After looking at the built-in datasets, I've decided that I want to use "ChickenWeight."

# Step 2: Before we begin any analyses, let's examine what this dataset includes:

```{r exploring chickens}
View(ChickWeight)
?ChickWeight
```

### The "view()" function opens the full dataset file in a new tab, so we can visually examine what the data looks like. The ?ChickenWeight function opens a help file that explains what this built-in dataset covers. 

### The help window tells us that "ChickenWeight" examines the weight versus age of chicken on different diets. This dataset contains 578 rows and 4 columns based on an experiment on the effect of diet on early growth of chicks. 

# Step 3: Let's explore some characteristics about the data:

```{r data characteristics}
glimpse(ChickWeight)
head(ChickWeight)
names(ChickWeight)
missing <- !complete.cases(ChickWeight)
ChickWeight[missing]
```

### - glimpse() shows us an overview of the whole dataset. We see that there are 578 rows and 4 columns, just like the help window told us.
### - head() gives us the first 6 rows of data
### - names() tells us all of the variable names: "weight" "Time" "Chick" "Diet"  
### - missing <- !complete.cases() creates an object/tibble called "missing" with any missing observations across the entire dataset
### - ChickWeight[missing] shows us that the "missing" tibble has 0 columns. This means that there was no missing data in this dataset, which makes it easier to work with! 

# Step 4: Let's get some basic descriptive statistics about each variable. 
