---
title: "Portfolio 2"
author: "Cassandra Richardson"
date: "4/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(tidyverse)
```

# **Cassandra Richardson's Portfolio Piece #2 - Data Analysis Report on Chicken Weight and Diet**



# Step 1: Finding a dataset.

### R has several built-in datasets that we can use to practice R functions with. To view these datasets, simply use the function below: 

```{r finding dataset}
data()
```

### After looking at the built-in datasets, I've decided that I want to use "ChickenWeight."

# Step 2: Before we begin any analyses, let's examine what this dataset includes:

```{r exploring chickens}
View(ChickWeight)
?ChickWeight
```

### The "view()" function opens the full dataset file in a new tab, so we can visually examine what the data looks like. The ?ChickenWeight function opens a help file that explains what this built-in dataset covers. 

### The help window tells us that "ChickenWeight" examines the weight versus age of chicken on different diets. This dataset contains 578 rows and 4 columns based on an experiment on the effect of diet on early growth of chicks. 

# Step 3: Let's explore some characteristics about the data:

```{r data characteristics}
glimpse(ChickWeight)
head(ChickWeight)
names(ChickWeight)
missing <- !complete.cases(ChickWeight)
ChickWeight[missing]
```

### - glimpse() shows us an overview of the whole dataset. We see that there are 578 rows and 4 columns, just like the help window told us.
### - head() gives us the first 6 rows of data
### - names() tells us all of the variable names: "weight" "Time" "Chick" "Diet"  
### - missing <- !complete.cases() creates an object/tibble called "missing" with any missing observations across the entire dataset
### - ChickWeight[missing] shows us that the "missing" tibble has 0 columns. This means that there was no missing data in this dataset, which makes it easier to work with! 

# Step 4: For the sake of practice, let's see what our data would look like in a wider format.

```{r pivot wide}
chickens_wide <- ChickWeight |> 
  pivot_wider(id_cols = c(Time, Chick), 
              names_from = Diet, 
              values_from = weight)
```

### In the code above, we did a univariate pivot to make our original dataset wider such that the column for diet (diet 1-4) was split into four columns for each type of diet, and weight was moved to the appropriate diet for each chick and time point. However, this format is harder to interpret and includes many missing variables. Moving forward, we won't use the "chicken_wide" data, and instead we will use the original "ChickWeight."

# Step 5: Let's get some basic descriptive statistics about each variable. 


```{r descriptives}
chicken_data <- ChickWeight

summary <- summary(chicken_data)

print(summary)
```

### The table printed above gives us a summary of descriptive statistics for all variables. 


# Step 6: Let's get the weight of the chickens for each of the four diets:

```{r chicken diets}
chicken_diets <- chicken_data |>
  group_by(Diet)|>
  summarize(
    mean = mean(weight),
    sd = sd(weight),
    min = min(weight),
    max = max(weight)
  )

print(chicken_diets)
```

### In the code above, I used dplyr's "group_by()" function to say that I want to group the descriptives into each of the four chicken diets. After the data is grouped by diet, I found the mean, sd, min, and max chicken weights for each diet. It looks like diet 1 had the lowest mean weight, whereas diet 3 had the highest mean weight. 

# Step 7: Let's make some plots to better visualize the chickens' weight by diet! 

```{r bar plots}


```
